Index measures for oak decline severity using phenotypic descriptors

### 2019-12-16 manuscript/manuscript.Rmd
finished introduction

Oak declines are complex syndromes that are facilitated by a variety of both abiotic and biotic factors.
Abiotic factors can include both predisposing and inciting factors such as drought and are often
Biotic agents can include insect infestation such as the larvae of bark boring beetle *Agrilus biguttatus* or bacterial stem bleeds associated with acute oak decline (AOD) [@brown2017; @denman2014].

@@

- Monitoring currently focusses on crown assessments

Machine learning is finding increasing utility in biological and geographical applications due to it's ability to learn complex patterns within and between variables.
A popular algorithm is that of random forest due to it's versatility, relatively trivial training requirements and resilience against over fitting [@breiman2001a] (refs).

- Random forest has previously been used in a forest condition context [@vitale2014].

It is an ensemble machine learning method where forests of bootstrap sampled decision trees are grown, yielding both out of bag sample prediction accuracies and variable importance.
It is also able to simultaneously handle both continuous and categorical variables.
Unsupervised learning is also possible [@shi2006]

- Unsupervised learning for complex patterns between descriptors.

@@

- Scalability of method for wider application
- Accessibility of collecting descriptors
- Discussion of applications to forestry monitoring and management
- Sensitivity for integration with molecular data compared to categorical assignments
- Sensitivity provided for integration with large scale omics analyses as well as long term monitoring for forest condition
- Refinement of protocol for large scale surveying

>>>>

Oak decline is becoming/predicted to become increasingly prevalent in the UK and across Europe with the onset of climate change (ref).
Oak declines are complex syndromes that are facilitated by a variety of both abiotic and biotic factors.
Abiotic factors can include both predisposing and inciting factors such as drought and are often
Biotic agents can include insect infestation such as the larvae of bark boring beetle *Agrilus biguttatus* or bacterial stem bleeds associated with acute oak decline (AOD) [@brown2017; @denman2014].

@@

Machine learning is finding increasing utility in ecological and environmental applications due to it's ability to learn complex patterns within and between variables [@cutler2007;@belgiu2016].
A popular algorithm is that of random forest due to it's versatility, relatively trivial training requirements and resilience against over fitting [@breiman2001a] (refs).

It is an ensemble machine learning method where forests of bootstrap sampled decision trees are grown, yielding both out of bag sample prediction accuracies and variable importance.
It is also able to simultaneously handle both continuous and categorical variables.
Unsupervised learning is also possible using the assumption that if the data contains any underlying structure, this should be distinguishable from a randomly generated version of itself [@shi2006].
An observation proximity matrix can be derived from this process on which additional unsupervised methods can be applied such as multidimensional scaling (ref).

@@

- Scalability of method for wider application
- Accessibility of collecting descriptors
- Discussion of applications to forestry monitoring and management
- Sensitivity for integration with molecular data compared to categorical assignments
- Sensitivity provided for integration with large scale omics analyses as well as long term monitoring for forest condition
- Refinement of protocol for large scale surveying
- Aid long term monitoring by providing comparibility over time. This was highlighted in the recently published Action Oak Knowledge Review as a key requirment in the future for safeguarding oak health. [@quine2019],
- Applied to long established monitoring data sets such as The International Co-operative Programme on Assessment and Monitoring of Air Pollution Effects on Forests (ICP Forests) [@lorenz1995].
